---
title: "Santander data exploration"
author: "V. Bezgachev"
date: "28 Oktober 2016"
output: html_document
---

# Introduction
Here I will explore the data provided by Santander for their Kaggle competition 
'Santander Product Recommendation'. We have two files with train and test data from 
them: 'train_ver2.csv' and 'test_ver2.csv'.

# Data description
Column Name     | Description
----------------|-------------
fecha_dato      | The table is partitioned for this column
ncodpers        | Customer code
ind_empleado    | Employee index: A active, B ex employed, F filial, N not employee, P pasive
pais_residencia | Customer's Country residence
sexo            | Customer's sex
age             | Age
fecha_alta      | The date in which the customer became as the first holder of a contract in the bank
ind_nuevo       | New customer Index. 1 if the customer registered in the last 6 months.
antiguedad      | Customer seniority (in months)
indrel          | 1 (First/Primary), 99 (Primary customer during the month but not at the end of the month)
ult_fec_cli_1t  | Last date as primary customer (if he isn't at the end of the month)
indrel_1mes     | Customer type at the beginning of the month, 1 (First/Primary customer), 2 (co-owner ), P (Potential), 3 (former primary), 4(former co-owner)
tiprel_1mes     | Customer relation type at the beginning of the month, A (active), I (inactive), P (former customer),R (Potential)
indresi         | Residence index (S (Yes) or N (No) if the residence country is the same than the bank country)
indext          | Foreigner index (S (Yes) or N (No) if the customer's birth country is different than the bank country)
conyuemp        | Spouse index. 1 if the customer is spouse of an employee
canal_entrada   | channel used by the customer to join
indfall         | Deceased index. N/S
tipodom         | Addres type. 1, primary address
cod_prov        | Province code (customer's address)
nomprov         | Province name
ind_actividad_cliente | Activity index (1, active customer; 0, inactive customer)
renta           | Gross income of the household
segmento        | segmentation: 01 - VIP, 02 - Individuals 03 - college graduated
ind_ahor_fin_ult1 | Saving Account
ind_aval_fin_ult1 | Guarantees
ind_cco_fin_ult1  | Current Accounts
ind_cder_fin_ult1 | Derivada Account
ind_cno_fin_ult1  | Payroll Account
ind_ctju_fin_ult1 | Junior Account
ind_ctma_fin_ult1 | Más particular Account
ind_ctop_fin_ult1 | particular Account
ind_ctpp_fin_ult1 | particular Plus Account
ind_deco_fin_ult1 | Short-term deposits
ind_deme_fin_ult1 | Medium-term deposits
ind_dela_fin_ult1 | Long-term deposits
ind_ecue_fin_ult1 | e-account
ind_fond_fin_ult1 | Funds
ind_hip_fin_ult1  | Mortgage
ind_plan_fin_ult1 | Pensions
ind_pres_fin_ult1 | Loans
ind_reca_fin_ult1 | Taxes
ind_tjcr_fin_ult1 | Credit Card
ind_valo_fin_ult1 | Securities
ind_viv_fin_ult1  | Home Account
ind_nomina_ult1   | Payroll
ind_nom_pens_ult1 | Pensions
ind_recibo_ult1   | Direct Debit

# Train data
``` {r echo = FALSE, message = FALSE, warning = FALSE, results = 'hide', cache = TRUE}
library(data.table)

selectRow <- c('ncodpers', 'fecha_dato',
               'ind_empleado', 'pais_residencia', 'sexo', 'age', 
               'fecha_alta', 'ind_nuevo', 'antiguedad', 'indrel',
               'ult_fec_cli_1t', 'indrel_1mes', 'tiprel_1mes', 'conyuemp',
               'canal_entrada', 'ind_actividad_cliente', 'renta')

# read the train data from the file
train <- fread('train_ver2.csv', sep = ',', na.strings = "NA", 
               stringsAsFactors = FALSE, select = selectRow)
```

## Remove suspicious NA's
After train data load and checking their summary ...
``` {r echo = FALSE}
# summary
summary(train)
```
... we see that there are exactly `r nrow(train[is.na(train$age)])` rows with NA's in some
important columns (age, customer seniority, activity index and some others).  
If we take a closer look to a couple of such records ...
``` {r echo = FALSE}
head(train[is.na(train$age)], 2)
tail(train[is.na(train$age)], 2)
```
... then we see that all other fields except the date and customer code are empty.
So we can remove them from the training set.
``` {r echo = FALSE, cache = TRUE}
train <- train[!is.na(train$age)]
```

## Residence country and age
Let's have a look to the customer's residence country ...
``` {r message = FALSE, warning = FALSE, echo = FALSE}
library(rworldmap)

country <- train[, .N, by = pais_residencia]
plotData <- joinCountryData2Map(dF = country, joinCode = "ISO2", 
                                nameJoinColumn = "pais_residencia", verbose=F)
mapCountryData(mapToPlot = plotData, nameColumnToPlot = "N", catMethod = "logFixedWidth",
                 oceanCol = "steelblue1",missingCountryCol = "white",
                 mapTitle = "Customers by country",
                 aspect = "variable")
```
... the most customers are coming from Spain. We will narrow down the data using the age 
limitations for Spain; this means, from 18 (age when any may open a bank account) to 
83 (life expectation). To be a bit more sure, I say that the valid age interval is from 18 to 90. 
For other ages I set a median value.

``` {r echo = FALSE}
train <- train[train$age >= 18 & train$age <= 83]
```

## Income
There are `r nrow(train[is.na(train$renta)])` rows where income is not available.
It is not possible to use decision trees to determine missing incomes since the rows
have new labels (country codes, some dates, etc.). For random forest it is necessary 
to have no NA's. So we take the median.

``` {r echo = FALSE}
train[is.na(train$renta)]$renta <- median(train[!is.na(train$renta)]$renta)
```